Lawvere-Tierney Sheafification in Homotopy Type Theory

### 2016-04-06 paper/paper_lipics.tex
finish

\let\oldthebibliography=\thebibliography
  \let\endoldthebibliography=\endthebibliography
  \renewenvironment{thebibliography}[1]{%
    \begin{oldthebibliography}{#1}%
      \setlength{\parskip}{0ex}%
      \setlength{\itemsep}{0ex}%
  }%
  {%
    \end{oldthebibliography}%
  }

@@

The idea is then to define sheaves in $\E$ as objects of $\E$ for
which it is impossible to make a distinction between objects and their
dense subobjects. This idea is formalized by saying that for every
dense subobject $A$ of $E$, the following canonical map is an
isomorphism
%
%\vspace{-0.7em}
\begin{equation}\label{equ:sheaf_def}
\Hom{\E}(E,F) \rightarrow \Hom{\E}(A,F).
%\vspace{-0.7em}
\end{equation}%
%
One can show that $\Sh{\E}$, the full sub-category of $\E$ given by
sheaves, is again a topos, with classifying object
$
\Omega_j = \{ P \in \Omega \ | \ j P  = P \}.
$
Thus, in case of the double negation, the resulting topos is boolean  
and admits classical reasoning.

@@

      \displaystyle{\modal_{n+1}(T)} \defeq  
      \displaystyle{\sum_{u:T \to \Type_n^\modal} \!\!\!\!\modal_{-1} 
      \left\|
      \sum_{a:T} u= (\lambda t,~\modal_n (a=t))
      \right\|}
    \end{array}
% \vspace{-0.3em}
\end{equation}
where $\Type_n^\modal$ is the type of modal types for $\modal_n$.
The rest of the paper is an inductive proof that this definition produces a
sequence of left-exact modalities.

@@

For instance, the notion of separated objects (which is a technical
notion used in the construction of the sheafification functor), \ie
objects for which the morphism
(\ref{equ:sheaf_def}) is just an embedding, is generalized in two
different ways.  
%
The subobjects to be considered are given by $n$-truncated homotopy fibers,
while the morphism (\ref{equ:sheaf_def}) is still required to be an
embedding.

@@

Following~\cite{sets_in_hott}, it is possible to show that, for any
homotopy level $n$ and any type $B$, $\Type_n$ classifies subobjects
of $B$ with $n$-truncated homotopy fibers in the sense that there is
an equivalence
%
\begin{mymath}[0em]
  \chi : \sum_{A:\Type} \sum_{f:A \to B} \prod_{b\in B}
\IsType n\
\fib{f}{b} \xrightarrow{\sim} 
 (B \to \Type_n)
\end{mymath}%
%
 such that the usual subobject
 classifier diagram (\cite[Theorem 4.8.4]{hottbook}) is a pullback.
Therefore, in our construction, we will represent a subobject of a
type $B$ with $n$-truncated homotopy fibers either as a map $f:A\to B$
such that $\IsType n \fib{f}{b}$, either as a map $B\to \Type_n$.
% \begin{mymath}
% \xymatrix{
%   A \ar[r]^{\hspace{-1em} t_f} \ar[d]_f & \Type_n^\bullet \ar[d]^{\pi_1}\\
%   B \ar[r]_{\hspace{-1em} \chi_f} & \Type_n
% }
% \end{mymath}
% is a pullback for any $f$ with
 % $n$-truncated homotopy fibers where $\Type_n^\bullet \defeq
 % \sum_{A:\Type_n} A$ is the universe of pointed
% $n$-truncated types and 
% \begin{mymath}t_f = \lambda
  % a,~(\fib{f}{f(a)},(a,\mathrm{idpath})).
% \end{mymath}

@@

Then $P$ is the {\em colimit} of $D$ if for any other $X:\Type$, 
\begin{mymath}[-0.5em]\IsEquiv\left(\lambda f:P \to X, \left( \lambda i,~f \circ q_i\, ;\, \lambda i\, j\,
  \phi,\, f (p_{i, j}^\phi)) \right)\right).\end{mymath}
\end{defi}
Using higher inductive types, every diagram $D$ on a graph $G$ admits a
colimit in homotopy type theory.

@@

% We use the following definition of truncated left exact modalities:
\begin{defi}
  \label{sec:defin-basic-prop-1}
  Let $n\geq -1$ be a truncation index. A left exact modality at level
  $n$ is the data of
  \begin{enumerate}[(i)]
  \item A predicate $P:\Type_n \to \HProp$
  \item For every $n$-truncated type $A$, a $n$-truncated type
    $\modal A$ such that $P(\modal A)$
  \item For every $n$-truncated type $A$, a map $\eta_A:A \to
    \modal A$
  \end{enumerate}
  such that
  \begin{enumerate}[(i)]
    \setcounter{enumi}{3}
  \item For every $n$-truncated types $A$ and $B$, if $P(B)$ then
    \begin{mymath}\left\{
      \begin{array}{rcl}
        (\modal A \to B) & \to & (A \to B) \\
        f & \mapsto & f \circ \eta_A
      \end{array} \right.\end{mymath}%
    is an equivalence.
  \item for any $A:\Type_n$ and $B:A \to \Type_n$ such that $P(A)$
    and $\prod_{x:A} P(B x)$, then $P\left( \sum_{x:A} B(x)\right)$
  \item for any $A:\Type_n$ and $x,y:A$, if $\modal A$ is
    contractible, then $\modal (x=y)$ is contractible.
  \end{enumerate}
  Conditions (i) to (iv) define a {\em reflective subuniverse}, (i) to
  (v) a {\em modality}.
\end{defi}

The type of all modal $n$-types, \ie types for which $P$ holds, is
defined as \footnote{Note that we omit the $n$
on $\modal_n$ in $\Type_n^\modal$ to ease the reading.} %$\Type^\modal_p$, so 
\begin{mymath}
\Type_n^\modal \ \defeq \sum_{T : \Type_n} P (T).
\end{mymath}%
%
Since basic type formers (dependent products, products, sigma types)
preserve truncation levels, we can leverage all theorems on modalities
presented in~\cite[Chapter 7.7]{hottbook}.
%
We now summarize the main properties that are used in our
formalization. We do not recall the proofs as they are not specific to
our presentation (also the setting using truncated types provides
minor changes), but they can be found in our Coq formalization (see
Section~\ref{sec:formalization}).

@@

As the definition of sheafification is based on subobject classifiers
and thus on homotopy fibers, we need to take a closer look at the way left
exact modalities preserve homotopy fibers. The following is a part of
the library~\cite{hottlib}.
%
\begin{prop}
\label{sec:defin-basic-prop}
For any $n$-truncated types $X$ and $Y$,
and any map $f:X \to Y$, the modalisation of fiber of $f$ above any element $y:Y$
is the fiber of $\modal f$ above $\eta_Y y$:
\begin{mymath}[-0.3em]\modal \left(\sum_{x:X}  (f(x) = y)\right) = \sum_{x:\modal X}
(\modal f(x) = \eta_Y(y)),\end{mymath}%
where $\modal f:\modal X \to \modal Y$ is defined via the equivalence {\it (iv)} and $\eta_Y$.

@@

@@

With topos theory point of view, it means that given a map $\sum_{e:E}
\chi\, e \to F$,
if there is an extension $\tilde f:E\to F$, then it is unique.
Actually, there is a simpler characterisation:

\begin{prop}
  A type $F$ is $\Type_{n+1}$ is separated if, and only if all its
  path types are $n$-modal, ie
  \begin{mymath}[-0.3em]\prod_{x,y:F}~\left( \modal_n(x=y) \right) = (x=y).\end{mymath}%
\end{prop}

@@

From any left exact modality $\modal_{-1}$ on $\HProp$, we define by
induction on the truncation index $n$, a
left exact modality $\modal_n$ at level $n$ that extends $\modal_{-1}$ in the
sense that the following property holds.
%
% We have to be careful: if we want to propagate the properties of the
% first modality $\modal_{-1}$, all the higher modalities $\modal_n$ must be compatible
% with $\modal_{-1}$, \ie we want the property
\begin{cond}\label{cond:hprop}
  For any mere proposition $P$ (where $\widehat P$ is $P$ seen as a
  $\Type_n$),  $\modal_n \widehat P = \modal_{-1} P$ and the
  following coherence diagram commutes
  \begin{mymath}\xymatrix{
    P \ar@{->}^{\sim}[r] \ar[d]_{\eta_{-1}} & \widehat P \ar[d]^{\eta_n} \\
    \modal_{-1} P \ar@{->}^{\sim}[r] & \modal_n \widehat P 
  }\end{mymath}%
\end{cond}

@@

As many proofs are very technical, we only give here sketches, and
often refer to formalization for all details.

@@

Let $T : \Type_{n+1}$. We define $\separated T$ as the image (see
Section~\ref{ssec:hott}) of
$\modal_n^T \circ \{\cdot\}_T$, as in
\begin{mymath}\xymatrix{
    T \ar[r]^{\{\cdot\}_T} \ar[d]_{\mu_T} & \left(\Type_n\right)^T \ar[d]^{\modal_n^T} \\
  \separated T \ar[r]& \left( \Type_n^\modal \right)^T
}, \end{mymath}%
where $\{\cdot\}_T$ is the singleton map $\lambda (t:T),~\lambda
(t':T),~t=t'$. 
%
$\separated T$ can be given explicitly by
%
\begin{mymath}
%\begin{array}{rclcl}
\separated T \defeq  \im (\lambda~t:T,~\lambda~ t',~ \modal_n (t = t')).
 %          \left( =  \sum\limits_{u:T \to \Type_n^\modal} \left\| \sum_{a:T} 
  %          (\lambda t,~\modal_n (a=t)) = u\right\|\right).
%\end{array}
\end{mymath}%
%
This corresponds to the free separated object used in the topos-theoretic construction, but using $\Type_n^\modal$ instead of the
$j$-subobject classifier $\Omega_j$.
%
\begin{prop}
  For any $T\!:\!\Type_{n+1}$, $\separated T$ is separated.  
\end{prop}

@@

Now, let $Q$ be any separated $\Type_{n+1}$, and $f:X \to Q$. Then this
 diagram commutes
\begin{mymath}[-0.5em]\xymatrix{
\|X\|_{n+1} \ar[r] \ar[rd]_{|f|_{n+1}} & \|\mathring T_{X}\|_{n+1} \ar[r] \ar[d] & \|\mathring
  T_{\mathring T_X} \|_{n+1} \ar[ld]\cdots \\
  & Q &
} \end{mymath}% 
% \begin{mymath}\xymatrix @C=4em{ 
%     \cdots \overline{\Delta_3} \tar[r]{4pt} \ar[rd]_{f\circ \pi_1} & \overline{\Delta_2}
%     \dar[r]{2pt} \ar[d]^{f\circ \pi_1} &
%     \overline{\Delta_1} = T \ar[ld]^{f\circ \pi_1}\\
%     & Q &
%   }\end{mymath}
But we know (lemma~\ref{lem:sepiscolim}) that $\separated T$ is the
$(n+1)$-colimit of the diagram $\mathring T$, thus there is an universal
arrow $\separated T \to Q$.
%
This is enough to prove that $(\separated,\mu)$ defines a reflective
subuniverse on $\Type_{n+1}$.
We can now state
\begin{prop}\label{prop:sep-mod}
  $(\separated,\mu)$ defines a modality on $\Type_{n+1}$.
\end{prop}

@@

%\nt{put the definition of $\modal_{n+1}$ upfront}
For any $T$ in $\Type_{n+1}$, 
$\modal_{n+1}T$ is defined as the closure of $\separated T$,
seen as a subobject of $T \to \Type_n^\modal$. 
%
$\modal_{n+1}T$ can be given explicitly by
\begin{mymath}
\modal_{n+1} T \ \defeq \sum_{u:T \to \Type_n^\modal} \modal_{-1}\left\| \sum_{a:T} 
            (\lambda t,~\modal_n (a=t)) = u\right\|.
\end{mymath}%

@@

As $T\to \Type_n^\modal$ is a sheaf, and $\modal_{n+1}T$ is closed in
$T\to \Type_n^\modal$, $\modal_{n+1}T$ is a sheaf. We now prove that
it forms a modality.

@@

A Coq formalization based on the Coq/HoTT library~\cite{hottlib} is available at
\url{https://github.com/KevinQuirin/sheafification}.
%
% After reviewing the content and some statistics about the
% formalization in Section~\ref{sec:cont-form}, we present the
% limitations of our formalization in Section~\ref{sec:limit-form}, in
% particular the issues relative to universe polymorphism.

\subsection{Content of the formalization}
\label{sec:cont-form}

% We provide a more detailed insight of the structure of our formalization:
% \begin{itemize} 
% \item Colimits and iterated kernel pairs are formalized in
% \texttt{Limit}, \texttt{T.v}, \texttt{OT.v}v \texttt{OT\_Tf.v}, \texttt{T\_telescope.v},\\ \texttt{Tf\_Omono\_sep.v}.% (552 lines).
% \item
% Reflective subuniverses and modalities are formalized in\\
% \texttt{reflective\_subuniverse.v}, \texttt{modalities.v}. % (1053 lines).
% \item 
% %
%   The definition of the dense topology as a left exact modality on
%   $\HProp$ is given in \texttt{sheaf\_base\_case.v}. % (186 lines).
% \item
% Section~\ref{sec:sheaves} is formalized in
% \texttt{sheaf\_def\_and\_thm.v}. % (1029 lines).
% \item
% Section~\ref{sec:sheafification} is formalized in
% \texttt{sheaf\_induction.v}. % (2340 lines).
% \end{itemize}

Overall, % with other files containing technical lemmas,
the project
contains 7914 lines, and it could be reduced a bit by improving the
way Coq tries to rewrite and apply lemmas automatically. 
The \texttt{coqwc} tool counts 1611 lines of specifications
(definitions, lemmas, theorems, propositions) and 5403 lines of proof
script.
%
This constitutes a significant amount of work but the part dedicated
to sheaves and sheafification is only 2200 lines of proof script,
which seems quite reasonable and encouraging, because it suggests
that homotopy type theory provides a convenient tool to formalize some
part of the theory of higher topoï. As a matter of comparison, this
\TeX\xspace file contains approximatively 2000 lines.

% In our formalization, some things are admitted, either because they
% are just an adaptation of a proposition already formalized
% in~\cite{hottbook}, either because they are true, but very technical:
% \begin{itemize}
% \item in \texttt{modalities.v}, an equality is admitted in the proof
%   of proposition~\ref{sec:defin-basic-prop}.
% \item in \texttt{nat\_lemmas.v}, the proof that $\IsHProp (n\leqslant
%   m)$ is admitted.
% \item in \texttt{sheaf\_induction.v}, in the proof that
%   $C(\mu_T)=(C\id)$ (see lemma~\ref{lem:sepiscolim}), we admit the proof
%   of equality of projections.
% \item in \texttt{sheaf\_induction.v}, we admit some very technical
%   proofs in proofs that $\square$ and $\star$ are modalities.
% \item in \texttt{sheaf\_induction.v}, cumulativity is an axiom,
%   because of universes issues.
% \end{itemize}

\subsection{Limitations of the formalization}
\label{sec:limit-form}

In the formalization, we had to use the \texttt{type-in-type} option, to handle
the universe issues we faced.

@@

%
This restriction in our formalization may be solved by
generalizing the management of universe polymorphism for recursive definition
%
or by the use of general ``resizing axiom'' which is still under
discussion in the community.

@@

As a future work, we would like to improve this construction in three
ways.
% \begin{itemize}
% \item 
(i) The extension to whole $\Type$ in lemma~\ref{lem:type} is not
  totally satisfactory, as every type is collapsed to a truncated
  one. But some types in homotopy type theory are not
  truncated~\cite[Example 8.8.6]{hottbook}.
  % At the moment, sheafification only works with truncated
  % types. This is fine from a programming point of view, 
  % but many important types in homotopy theory are not
  % truncated, \eg the circle $\mathbb S^2$.  Thus, we plan to extend
  % the sheafification to all types by using a
  % limit construction.
% \item 
(ii) 
  We would like to have more examples of left-exact modalities on
  $\HProp$, in order to extend type theory with different principles
  than excluded middle.
% \item 
% (iii) It would be useful to improve the translation induced by the
%   sheafification modality to preserve more definitional equalities.
%   %
%   This could be done for instance by defining a higher inductive type
%   isomorphic to the modality but with a better computational content. 
  % able to work in the new type theory into a proof assistant.
% \end{itemize}

Moreover, in topos theory, Lawvere-Tierney subsumes Grothen\-dieck~\cite[Section~V.4]{maclanemoerdijk} in the sense that any
Grothen\-dieck topology gives rise to a Lawvere-Tierney topology with
the same notion of sheaves. Higher Lawvere-Tierney sheaves are
presented here, and higher Grothendieck sheaves have been defined
in~\cite{lurie}. It should be interesting to check if the subsumption
remains true in higher topos theory.

@@

%\bibliographystyle{abbrvnat}
\bibliography{sheaf_lics}

\newpage
\section*{Appendix}

@@

\subsection{Translation}
\label{ssec:translation}

>>>>

% \let\oldthebibliography=\thebibliography
%   \let\endoldthebibliography=\endthebibliography
%   \renewenvironment{thebibliography}[1]{%
%     \begin{oldthebibliography}{#1}%
%       \setlength{\parskip}{0ex}%
%       \setlength{\itemsep}{0ex}%
%   }%
%   {%
%     \end{oldthebibliography}%
%   }

@@

The idea is then to define sheaves in $\E$ as objects of $\E$ for
which it is impossible to make a distinction between objects and their
dense subobjects. This idea is formalized by saying that for every
dense subobject $A$ of $E$, the canonical map $\Hom{\E}(E,F) \rightarrow \Hom{\E}(A,F)$ is an
isomorphism.
%
%\begin{equation}\label{equ:sheaf_def}
%\Hom{\E}(E,F) \rightarrow \Hom{\E}(A,F).
%\end{equation}%
%
One can show that $\Sh{\E}$, the full sub-category of $\E$ given by
sheaves, is again a topos, with classifying object
$
\Omega_j = \{ P \in \Omega \ | \ j P  = P \}.
$
Thus, in case of the double negation, the resulting topos is boolean  
and admits classical reasoning.

@@

      \displaystyle{\modal_{n+1}(T)} \defeq  
      \displaystyle{\sum_{u:T \to \Type_n^\modal} \!\!\!\!\modal_{-1} 
      \left\|
      \sum_{a:T} u= (\lambda t,~\modal_n (a=t))
      \right\|}
    \end{array}
\end{equation}
where $\Type_n^\modal$ is the type of modal types for $\modal_n$.
The rest of the paper is an inductive proof that this definition produces a
sequence of left-exact modalities.

@@

For instance, in the notion of separated objects defined in section~\ref{sec:sheafification},
%(which is a technical
%notion used in the construction of the sheafification functor), \ie
%objects for which the morphism
%(\ref{equ:sheaf_def}) is just an embedding, 
subobjects are generalized in two
different ways.  
%
The subobjects to be considered are given by $n$-truncated homotopy fibers,
while the morphism $\Hom{\E}(E,F) \rightarrow \Hom{\E}(A,F)$ is still required to be an
embedding.

@@

Following~\cite{sets_in_hott}, it is possible to show that, for any
homotopy level $n$ and any type $B$, $\Type_n$ classifies subobjects
of $B$ with $n$-truncated homotopy fibers in the sense that there is
an equivalence
%
\begin{mymath}[-0em]
  \chi : \sum_{A:\Type} \sum_{f:A \to B} \prod_{b\in B}
\IsType n\
\fib{f}{b} \xrightarrow{\sim} 
 (B \to \Type_n)
\end{mymath}%
%
 such that the usual subobject
 classifier diagram (\cite[Theorem 4.8.4]{hottbook}) is a pullback.
Therefore, in our construction, we will represent a subobject of a
type $B$ with $n$-truncated homotopy fibers either as a map $f:A\to B$
such that $\IsType n \fib{f}{b}$, either as a map $B\to \Type_n$.
% \begin{mymath}
% \xymatrix{
%   A \ar[r]^{\hspace{-1em} t_f} \ar[d]_f & \Type_n^\bullet \ar[d]^{\pi_1}\\
%   B \ar[r]_{\hspace{-1em} \chi_f} & \Type_n
% }
% \end{mymath}
% is a pullback for any $f$ with
 % $n$-truncated homotopy fibers where $\Type_n^\bullet \defeq
 % \sum_{A:\Type_n} A$ is the universe of pointed
% $n$-truncated types and 
% \begin{mymath}t_f = \lambda
  % a,~(\fib{f}{f(a)},(a,\mathrm{idpath})).
% \end{mymath}

@@

Then $P$ is the {\em colimit} of $D$ if for any other $X:\Type$, 
\begin{mymath}[-0em]\IsEquiv\left(\lambda f:P \to X, \left( \lambda i,~f \circ q_i\, ;\, \lambda i\, j\,
  \phi,\, f (p_{i, j}^\phi)) \right)\right).\end{mymath}
\end{defi}
Using higher inductive types, every diagram $D$ on a graph $G$ admits a
colimit in homotopy type theory.

@@

% We use the following definition of truncated left exact modalities:
\begin{defi}
  \label{sec:defin-basic-prop-1}
  Let $n\geq -1$ be a truncation index. A left exact modality at level
  $n$ is the data of
  \begin{enumerate}[(i)]
  \item A predicate $P:\Type_n \to \HProp$
  \item For every $n$-truncated type $A$, a $n$-truncated type
    $\modal A$ such that $P(\modal A)$
  \item For every $n$-truncated type $A$, a map $\eta_A:A \to
    \modal A$
  \end{enumerate}
  such that
  \begin{enumerate}[(i)]
    \setcounter{enumi}{3}
  \item For every $n$-truncated types $A$ and $B$, if $P(B)$ then
    \begin{mymath}[-0em]\left\{
      \begin{array}{rcl}
        (\modal A \to B) & \to & (A \to B) \\
        f & \mapsto & f \circ \eta_A
      \end{array} \right.\end{mymath}%
    is an equivalence.
  \item for any $A:\Type_n$ and $B:A \to \Type_n$ such that $P(A)$
    and $\prod_{x:A} P(B x)$, then $P\left( \sum_{x:A} B(x)\right)$
  \item for any $A:\Type_n$ and $x,y:A$, if $\modal A$ is
    contractible, then $\modal (x=y)$ is contractible.
  \end{enumerate}
  Conditions (i) to (iv) define a {\em reflective subuniverse}, (i) to
  (v) a {\em modality}.
\end{defi}

The type of all modal $n$-types, \ie types for which $P$ holds, is
defined as \footnote{Note that we omit the $n$
on $\modal_n$ in $\Type_n^\modal$ to ease the reading.} %$\Type^\modal_p$, so 
\begin{mymath}[-0em]
\Type_n^\modal \ \defeq \sum_{T : \Type_n} P (T).
\end{mymath}%
%
Since basic type formers (dependent products, products, sigma types)
preserve truncation levels, we can leverage all theorems on modalities
presented in~\cite[Chapter 7.7]{hottbook}.
%
We now summarize the main properties that are used in our
formalization. We do not recall the proofs as they are not specific to
our presentation (also the setting using truncated types provides
minor changes), but they can be found in our Coq formalization (see
Section~\ref{sec:formalization}).

@@

As the definition of sheafification is based on subobject classifiers
and thus on homotopy fibers, we need to take a closer look at the way left
exact modalities preserve homotopy fibers. The following is a part of
the library~\cite{hottlib}.
%
\begin{prop}
\label{sec:defin-basic-prop}
For any $n$-truncated types $X$ and $Y$,
and any map $f:X \to Y$, the modalisation of fiber of $f$ above any element $y:Y$
is the fiber of $\modal f$ above $\eta_Y y$:
\begin{mymath}[-0em]\modal \left(\sum_{x:X}  (f(x) = y)\right) = \sum_{x:\modal X}
(\modal f(x) = \eta_Y(y)),\end{mymath}%
where $\modal f:\modal X \to \modal Y$ is defined via the equivalence {\it (iv)} and $\eta_Y$.

@@

For space reason, the main ideas of the proofs are given in the
appendix, and we refer the reader to the formalization for all details.

@@

With topos theory point of view, it means that given a map $\sum_{e:E}
\chi\, e \to F$,
if there is an extension $\tilde f:E\to F$, then it is unique.
Equivalently, it can be checked that a type is separated if, and only if all its path types
are $n$-modal.

% \begin{prop}
%   A type $F$ is $\Type_{n+1}$ is separated if, and only if all its
%   path types are $n$-modal, ie
%   \begin{mymath}[-0.3em]\prod_{x,y:F}~\left( \modal_n(x=y) \right) = (x=y).\end{mymath}%
% \end{prop}

@@

From any left exact modality $\modal_{-1}$ on $\HProp$, we define by
induction on the truncation index $n$, a
left exact modality $\modal_n$ at level $n$ that extends $\modal_{-1}$ in the
sense that the following property holds.
%
% We have to be careful: if we want to propagate the properties of the
% first modality $\modal_{-1}$, all the higher modalities $\modal_n$ must be compatible
% with $\modal_{-1}$, \ie we want the property
\begin{cond}\label{cond:hprop}
  For any mere proposition $P$ (where $\widehat P$ is $P$ seen as a
  $\Type_n$),  $\modal_n \widehat P \simeq \modal_{-1} P$ and the
  following coherence diagram commutes
  \begin{mymath}[-0em]\xymatrix{
    P \ar@{->}^{\sim}[r] \ar[d]_{\eta_{-1}} & \widehat P \ar[d]^{\eta_n} \\
    \modal_{-1} P \ar@{->}^{\sim}[r] & \modal_n \widehat P 
  }\end{mymath}%
\end{cond}

@@

@@

Let $T : \Type_{n+1}$. We define $\separated T$ as the image (see
Section~\ref{ssec:hott}) of
$\modal_n^T \circ \{\cdot\}_T$, as in
\begin{mymath}[-0em]\xymatrix{
    T \ar[r]^{\{\cdot\}_T} \ar[d]_{\mu_T} & \left(\Type_n\right)^T \ar[d]^{\modal_n^T} \\
  \separated T \ar[r]& \left( \Type_n^\modal \right)^T
}, \end{mymath}%
where $\{\cdot\}_T$ is the singleton map $\lambda (t:T),~\lambda
(t':T),~t=t'$. 
%
$\separated T$ can be given explicitly by
%
\begin{mymath}[-0em]
%\begin{array}{rclcl}
\separated T \defeq  \im (\lambda~t:T,~\lambda~ t',~ \modal_n (t = t')).
 %          \left( =  \sum\limits_{u:T \to \Type_n^\modal} \left\| \sum_{a:T} 
  %          (\lambda t,~\modal_n (a=t)) = u\right\|\right).
%\end{array}
\end{mymath}%
%
This corresponds to the free separated object used in the topos-theoretic construction, but using $\Type_n^\modal$ instead of the
$j$-subobject classifier $\Omega_j$.
%
\begin{prop}
  For any $T\!:\!\Type_{n+1}$, $\separated T$ is separated.  
\end{prop}

@@

Now, let $Q$ be any separated $\Type_{n+1}$, and $f:X \to Q$. 
Then $|f|_{n+1}$ can be extended to build a cocone over $\mathring
T$ into $Q$.

% Then this
%  diagram commutes
% \begin{mymath}[-0.5em]\xymatrix{
% \|X\|_{n+1} \ar[r] \ar[rd]_{|f|_{n+1}} & \|\mathring T_{X}\|_{n+1} \ar[r] \ar[d] & \|\mathring
%   T_{\mathring T_X} \|_{n+1} \ar[ld]\cdots \\
%   & Q &
% } \end{mymath}% 
% \begin{mymath}\xymatrix @C=4em{ 
%     \cdots \overline{\Delta_3} \tar[r]{4pt} \ar[rd]_{f\circ \pi_1} & \overline{\Delta_2}
%     \dar[r]{2pt} \ar[d]^{f\circ \pi_1} &
%     \overline{\Delta_1} = T \ar[ld]^{f\circ \pi_1}\\
%     & Q &
%   }\end{mymath}
But we know (lemma~\ref{lem:sepiscolim}) that $\separated T$ is the
$(n+1)$-colimit of the diagram $\mathring T$, thus there is an universal
arrow $\separated T \to Q$.
%
This is enough to prove that $(\separated,\mu)$ defines a reflective
subuniverse on $\Type_{n+1}$.
As $(\separated,\mu)$ is a modality and closure is universal, we can state
\begin{prop}\label{prop:sep-mod}
  $(\separated,\mu)$ defines a modality on $\Type_{n+1}$.
\end{prop}

@@

%\nt{put the definition of $\modal_{n+1}$ upfront}
For any $T$ in $\Type_{n+1}$, 
$\modal_{n+1}T$ is defined as the closure of $\separated T$,
seen as a subobject of $T \to \Type_n^\modal$. 
%
$\modal_{n+1}T$ can be given explicitly by
\begin{mymath}[-0em]
\modal_{n+1} T \ \defeq \sum_{u:T \to \Type_n^\modal} \modal_{-1}\left\| \sum_{a:T} 
            (\lambda t,~\modal_n (a=t)) = u\right\|.
\end{mymath}%

@@

As $T\to \Type_n^\modal$ is a sheaf, and $\modal_{n+1}T$ is closed in
$T\to \Type_n^\modal$, $\modal_{n+1}T$ is a sheaf. We can now state:

@@

A Coq formalization based on the Coq/HoTT library~\cite{hottlib} is available at
\url{https://github.com/KevinQuirin/sheafification}. This development
contains 7914 lines, with approximatively 20\% lines of
definitions, which seems quite reasonable and encouraging, because it
suggests that homotopy type theory provides a convenient tool to
formalize some part of the theory of higher topoï.
%
% After reviewing the content and some statistics about the
% formalization in Section~\ref{sec:cont-form}, we present the
% limitations of our formalization in Section~\ref{sec:limit-form}, in
% particular the issues relative to universe polymorphism.

% \subsection{Content of the formalization}
% \label{sec:cont-form}

% % We provide a more detailed insight of the structure of our formalization:
% % \begin{itemize} 
% % \item Colimits and iterated kernel pairs are formalized in
% % \texttt{Limit}, \texttt{T.v}, \texttt{OT.v}v \texttt{OT\_Tf.v}, \texttt{T\_telescope.v},\\ \texttt{Tf\_Omono\_sep.v}.% (552 lines).
% % \item
% % Reflective subuniverses and modalities are formalized in\\
% % \texttt{reflective\_subuniverse.v}, \texttt{modalities.v}. % (1053 lines).
% % \item 
% % %
% %   The definition of the dense topology as a left exact modality on
% %   $\HProp$ is given in \texttt{sheaf\_base\_case.v}. % (186 lines).
% % \item
% % Section~\ref{sec:sheaves} is formalized in
% % \texttt{sheaf\_def\_and\_thm.v}. % (1029 lines).
% % \item
% % Section~\ref{sec:sheafification} is formalized in
% % \texttt{sheaf\_induction.v}. % (2340 lines).
% % \end{itemize}

% Overall, % with other files containing technical lemmas,
% the project
% contains 7914 lines, and it could be reduced a bit by improving the
% way Coq tries to rewrite and apply lemmas automatically. 
% The \texttt{coqwc} tool counts 1611 lines of specifications
% (definitions, lemmas, theorems, propositions) and 5403 lines of proof
% script.
% %
% This constitutes a significant amount of work but the part dedicated
% to sheaves and sheafification is only 2200 lines of proof script,
% which seems quite reasonable and encouraging, because it suggests
% that homotopy type theory provides a convenient tool to formalize some
% part of the theory of higher topoï. As a matter of comparison, this
% \TeX\xspace file contains approximatively 2000 lines.

% % In our formalization, some things are admitted, either because they
% % are just an adaptation of a proposition already formalized
% % in~\cite{hottbook}, either because they are true, but very technical:
% % \begin{itemize}
% % \item in \texttt{modalities.v}, an equality is admitted in the proof
% %   of proposition~\ref{sec:defin-basic-prop}.
% % \item in \texttt{nat\_lemmas.v}, the proof that $\IsHProp (n\leqslant
% %   m)$ is admitted.
% % \item in \texttt{sheaf\_induction.v}, in the proof that
% %   $C(\mu_T)=(C\id)$ (see lemma~\ref{lem:sepiscolim}), we admit the proof
% %   of equality of projections.
% % \item in \texttt{sheaf\_induction.v}, we admit some very technical
% %   proofs in proofs that $\square$ and $\star$ are modalities.
% % \item in \texttt{sheaf\_induction.v}, cumulativity is an axiom,
% %   because of universes issues.
% % \end{itemize}

\paragraph*{On the use of universe polymorphism}
%\subsection{Limitations of the formalization}
%\label{sec:limit-form}

The development uses the new universe polymorphism feature of
Coq~\cite{sozeau2014universe}. However, it still lacks some features, and
in the formalization, we had to use the \texttt{type-in-type} option, to handle
the universe issues we faced.

@@

%
This restriction in our formalization may be solved by
generalizing the management of universe polymorphism for recursive definition
%
and by the use of general ``resizing axiom'' which is still under
discussion in the community.
%
We note that for this solution to work, we need to have a more general 
definition of modalities than the one in~\cite{hottlib}, which ask the
reflector $\modal$ to map a universe $\Type^i$ to
itself. Here, we relax this constraint by allowing the reflector to
map a universe $\Type^i$ to a
possibly higher universe $\Type^j$. 
We have checked that the library~\cite{hottlib} still works with this
generalized definition (see
\url{https://github.com/KevinQuirin/HoTT/tree/extended_modalities/} for the
formalization).

@@

As a future work, we would like to improve this construction in three
ways.
% \begin{itemize}
% \item 
(i) The extension to whole $\Type$ in lemma~\ref{lem:type} is not
  totally satisfactory, as every type is collapsed to a truncated
  one. But some types in homotopy type theory are not
  truncated~\cite[Example 8.8.6]{hottbook}.
  % At the moment, sheafification only works with truncated
  % types. This is fine from a programming point of view, 
  % but many important types in homotopy theory are not
  % truncated, \eg the circle $\mathbb S^2$.  Thus, we plan to extend
  % the sheafification to all types by using a
  % limit construction.
% \item 
(ii) 
  We would like to have more examples of left-exact modalities on
  $\HProp$, in order to extend type theory with different principles
  than excluded middle.
% \item 
% (iii) It would be useful to improve the translation induced by the
%   sheafification modality to preserve more definitional equalities.
%   %
%   This could be done for instance by defining a higher inductive type
%   isomorphic to the modality but with a better computational content. 
  % able to work in the new type theory into a proof assistant.
% \end{itemize}
%
(iii) In topos theory, Lawvere-Tierney subsumes Grothen\-dieck~\cite[Section~V.4]{maclanemoerdijk} in the sense that any
Grothen\-dieck topology gives rise to a Lawvere-Tierney topology with
the same notion of sheaves. Higher Lawvere-Tierney sheaves are
presented here, and higher Grothendieck sheaves have been defined
in~\cite{lurie}. It should be interesting to check if the subsumption
remains true in higher topos theory.

@@

%\bibliographystyle{abbrvnat}
\newpage
\bibliography{sheaf_lics}

\newpage
\appendix
\section{Proofs}

@@

\section{Translation}
\label{ssec:translation}
